% rubber: setlist arguments --shell-escape

\documentclass[
    fontsize=12pt,
    headings=small,
    parskip=half,
    bibliography=totoc,
    numbers=noenddot,
    open=any
]{scrreprt}

\usepackage[utf8]{inputenc}
\usepackage{datetime2}
\DTMlangsetup[en-US]{abbr}

\usepackage[titletoc,title]{appendix}

\usepackage[
    autostyle,
    ]{csquotes}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{pifont}
\usepackage[table]{xcolor}
\usepackage{marvosym}
\usepackage{graphicx}
\usepackage{parskip}
\usepackage{hyperref}
\usepackage{setspace}
\usepackage{acronym}
% \usepackage{float}
\usepackage[scaled=.92]{helvet}
\usepackage{courier}
\usepackage[
    a4paper,
    margin=2.54cm,
    marginparwidth=2.0cm,
    footskip=1.0cm
]{geometry}

\usepackage[
    url=true,
    style=numeric,
    alldates=long,
    dateabbrev=false
]{biblatex}

\usepackage{minted}

\addbibresource{literature.bib}

% \pagestyle{plain}
\urlstyle{rm}

\definecolor{hlinegray}{gray}{0.8}

\newcommand{\cmark}{\textcolor{green}{\ding{51}}}
% \newcommand{\xmark}{\textcolor{red}{\ding{55}}}
\newcommand{\xmark}{\textcolor{red}{-}}
\newcommand{\qmark}{\textcolor{orange}{\textbf{?}}}

\title{Data}

\author{Tronje Krabbe}

\date{\today}

\begin{document}
\hypersetup{hidelinks}

\setcounter{tocdepth}{1}
\tableofcontents

\chapter{Data Acquisition and Analysis}
This chapter presents the dataset used to conduct the experiments used in this thesis. We will describe
how the data was acquired, as well as some basic properties of the set.

\section{Acquiring Site Information}
As previously mentioned, we used OpenWPM\footnote{\url{https://github.com/citp/OpenWPM}} to gather data about
website behavior. OpenWPM visits websites with a Firefox browser instance, and records all JavaScript accesses to
attributes that may be used for fingerprinting, as well as other information, like HTTP headers, etc.

The parameters used with OpenWPM to gather the data can be found in Appendix~\ref{app:params}.

\subsection{Alex Top One Million List}
Alex regularly compiles a list of the top one million websites, ordered by the number of
visitors\footnote{\url{https://www.alexa.com/topsites}}.
We have analyzed the top 5000 websites, as described by a version of this list acquired from
\url{http://s3.amazonaws.com/alexa-static/top-1m.csv.zip} on December 19, 2017.


\subsection{Datasets}
Multiple datasets were gathered, all for the 5000 sites as described previously. Due to timeouts, or other
errors, not all 5000 sites were actually crawled each time.
The primary dataset we used was crawled on January 27, 2018, and includes 4929 sites with distinct URLs. Some of these sites
may be duplicates of each other. For instance, the URL \url{https://fbcdn.net/} for Facebook's content delivery
network simply points to \url{https://facebook.com/}, both of which are included in the dataset.


\section{Data Properties}
In this section, we will give an overview of basic properties of the primary dataset, with special regard to
their relevance for fingerprinting.

\subsection{Overview}
The table below shows some interesting, basic properties of our primary dataset. Percentages have been rounded.

\definecolor{hlinegray}{gray}{0.8}
\begin{center}
\begin{tabular}{l r r}
\arrayrulecolor{hlinegray}
    & amount & percentage \\
    \hline
    All sites & 4929 & 100\% \\
    \hline
    \\
    \textit{basic attributes} & & \\
    \hline
    \texttt{userAgent} lookup & 4424 & 90\% \\ %89.8
    \hline
    \texttt{language} or \texttt{languages} lookup & 3985 & 81\% \\ %80.8
    \hline
    \texttt{colorDepth} lookup & 3931 & 80\% \\ %79.8
    \hline
    \texttt{localStorage} lookup & 3772 & 77\% \\
    \hline
    \texttt{platform} lookup & 3275 & 66\% \\ %66.4
    \hline
    \texttt{sessionStorage} lookup & 2860 & 58\% \\
    \hline
    \texttt{cookieEnabled} lookup & 2823 & 57\% \\ %57.3
    \hline
    \texttt{doNotTrack} lookup & 1406 & 29\% \\ %28.5
    \hline
    \texttt{oscpu} lookup & 580 & 12\% \\ %11.8
    \hline
    \\
    \textit{using multiple of above attributes} & & \\
    \hline
    Sites using at least 2 attributes & 4338 & 88\% \\
    \hline
    Sites using at least 3 attributes & 4245 & 86\% \\
    \hline
    Sites using at least 4 attributes & 3968 & 81\% \\
    \hline
    Sites using at least 5 attributes & 3572 & 72\% \\
    \hline
    Sites using at least 6 attributes & 2906 & 59\% \\
    \hline
    Sites using at least 7 attributes & 1963 & 40\% \\
    \hline
    Sites using at least 8 attributes & 1138 & 23\% \\
    \hline
    Sites using 9 attributes & 473 & 10\% \\
    \hline
    \\
    Calling functions defined by \textit{Fingerprintjs2} & 336 & 7\% \\ %6.8
    \hline
    \\
    \textit{Canvas-related functions or attributes} & & \\
    \hline
    Using any canvas-related JavaScript & 1657 & 34\% \\ %33.6
    \hline
    \texttt{toDataURL or getImageData} & 884 & 18\% \\ %17.9
    \hline
    \texttt{toDataURL} & 768 & 16\% \\ %15.6
    \hline
    \texttt{getImageData} & 252 & 5\% \\ %5.1
    \hline
    \\
\end{tabular}
\end{center}

Because of the basic attributes' ease of access, we may hope that these attributes are consistently used by different fingerprinters.
Should that be the case, we can eliminate a large amount of sites from our dataset as non-fingerprinters;
if we, for instance, find out that every fingerprinting script uses the \texttt{window.navigator.doNotTrack}
property, 79\% of our crawled sites can be eliminated immediately.

%% obviously, as it is the 6th most common, and this discusses 6-attr sites...
% Indeed, as the table shows, 1216 sites use at least six of the described attributes. One of these six out of seven
% attributes was the \texttt{doNotTrack} attribute for 1178 sites; about 97\% of the six-attribute sites.

Of the 884 sites that use \texttt{toDataURL} or \texttt{getImageData} on a HTML5 Canvas object, which suggests
a high likelihood of canvas fingerprinting, 536, or about 61\%, also use eight or more of the basic attributes.
These 536 sites are good candidates for likely fingerprinters, and indeed, 147 of them overlap with the 336 sites that
we have identified as users of \textit{Fingerprintjs2}.


\begin{appendices}

\chapter{OpenWPM Browser Params}
\label{app:params}
The following configuration for OpenWPM's \textit{default\_browser\_params} was used.

\definecolor{mintedbg}{rgb}{0.95,0.95,0.95}
\setminted{fontsize=\footnotesize,bgcolor=mintedbg}
\begin{minted}{json}
{
    "extension_enabled": true,
    "cookie_instrument": false,
    "js_instrument": true,
    "cp_instrument": true,
    "http_instrument": true,
    "save_javascript": true,
    "save_all_content": false,

    "random_attributes": false,
    "bot_mitigation": false,
    "disable_flash": true,
    "profile_tar": null,
    "profile_archive_dir": null,
    "headless": true,
    "browser": "firefox",
    "prefs": {},

    "tp_cookies": "always",
    "donottrack": false,
    "disconnect": false,
    "ghostery": false,
    "https-everywhere": false,
    "adblock-plus": false,
    "ublock-origin": false,
    "tracking-protection": false
}
\end{minted}


\end{appendices}

\clearpage

\printbibliography

\end{document}
